### 2025-04-08 Paper.tex
Update Paper.tex

@@

>>>>

 	Our work presents a method for emulating the vocal style of the late Chester Bennington, the former lead singer of Linkin Park. Our goal was to synthesize a new version of a song from the latest album “From Zero” using Chester Bennington’s distinctive voice. For this purpose, we used a pre-trained Singing Voice Synthesis (SVS) model called Diffsinger.
We compiled a dataset from the complete official discography of Linkin Park, including both studio and live recordings. This mixed dataset was deliberately chosen to expose the model to a range of different voice variations and production techniques over the years in order to make the model more robust.
After obtaining the data, we proceeded to isolate the vocal tracks with the help of several deep learning-based source separation tools. Subsequently, the extracted vocal tracks were segmented and aligned in preparation for the training. In the end, we used Reaper to mix the model’s output and combined it with the instrumental track of the album.

@@

 	Mixing & Mastering

	The vocal track generated by DiffSinger posed some challenges during mixing. Inconsistent tones, robotic artifacts and a lack of natural dynamics all required extensive processing to attempt to improve its quality. Indeed, the output track contains glitches, sudden pitch shifts, muffled consonants and irregular sustain. The sustained notes sound unnaturally flat and lack vibrato. To address this, we first tried to trim segments and applied noise reduction using ReaFIR. A high-pass filter at 120 Hz (ReaEQ) removed the jarring low-frequency rumble. Given the irregular dB dynamics where some words were overly loud and others nearly inaudible. We automated a +2 dB rise during the chorus to compensate for the model’s weak projection on longer notes. To make up for some synthetic artifacts, we applied a cut at 2–3 kHz (-5 dB, Q=4) to reduce the volume of some harsh resonances. We then added a high-shelf boost at 8 kHz (+2 dB) for added clarity to handle occasional muffled sibilants. Moreover, compression was used to stabilize the uneven output and an even dB level. ReaComp set at 4:1 ratio, 20 ms attack, 100 ms release provided 6 dB of gain reduction to lower some peaks. We also used  a limiter (ReaLimit) to prevent the track from clipping. To account for the “s” and “t” sounds, we used the FabFilter PRO-DS vst-plugin. However, despite these efforts, the vocals retained an unnatural texture, which we attempted to mitigate, by applying some saturation, not just to add warmth, but also to Emulate Emily Armstrong’s raspy tone that the model failed to emulate. 
	Mastering
	In music production, mastering is the final stage before the release of a song or an album. The process consists of mixing and polishing the master track that comprises all single tracks in order to achieve consistency, optimal loudness and compatibility across various playback systems. Unlike mixing, which focuses on balancing individual tracks (vocals, instruments), mastering addresses the entire stereo mix as a single unite. Indeed, this phase aimed to unify the mix’s tonal balance and loudness. Thus, some EQ (FabFilter Pro-Q3) was applied to the stereo bus to mitigate the harshness at 3.5 kHz (-1.5 dB, Q=2) and add a touch of higher frequency with a high-shelf boost at 12 kHz (+1 dB). Multiband compression (FabFilter - MB) was applied to tighten the low-midrange (200–500 Hz) and reduce the “muddiness” that stems from the vocal’s uneven low-end resonance. Finally a limiter (FabFilter-L2) was set to +9 to boost loudness in accordance with modern rock and metal commercial standards.
